---
title: "OIP Postprocess Part 1"
author: "Paul Leiby and Rocio Uria"
date: "10/1/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(readxl)
```

## Data Section

#### Expected location of input data (result files from past runs of OIP)

Locations of input and outputs are all relative to location of project directory.

```{r}
# Warning: following is machine-dependent path
#  Instead, prefer to use here("..")
# machine_base_path = "/Users/paulleiby/Documents" # PL Mac
# machine_base_path = "D:Workfiles" # PL Amanda Home Desktop
# OIP_model_output_path = "/Papers/2006OilImportPremium/Analysis"

# Expected location of output files for each run of OIP (xlsx format)
# Warning: this is the path _relative_ to the project, and can look outside the project folder for data
# inputdata_subpath = "Data/inputdata" # path relative to project folder
inputdata_subpath <- ".." # OIP_model_output_path = here(".."), just above project dir location. Does this work

output_subpath <- "Output" # destination for any output

# Set logical flag for whether to save process results or not
save_results_to_CSV <- T
```

#### Specify set of OIP Runs to be Processed

```{r}
model_versionyears <- c("OIP2022", "OIP2023")
model_subversions <- c("v31", "v32") # each goes with the corresponding model_versionyear

model_versionyear <- model_versionyears[2]
model_subversion <- model_subversions[2]
```


```{r}
runsets <- paste0("r0", 1:4)
runsets
```

#### Following data specify one standard set of runs expected

More generally, we can just read whatever OIP run result files are found.

Standard OIP run result files have following name/location conventions:

- Begin with `model_versionyear`, `model_subversion`, run set indicator, and underscore, as in "OIPYYYY" "v31" "r01_", i.e. `r paste0(model_versionyear, model_subversion, paste0("r0", 1))`
- Are Excel data files ending with extension ".xlsx"
- Are stored in neighboring directory `toupper(model_subversion)`, as in `r toupper(model_subversion)`
- Each set is stored in subdirectory bases on `model_subversion` and `run_num`, as in, for `runset_num` 1: `r runsets[1]`


```{r}
# follopwing is a particular set of years that may be evaluated, 
# and a default string of switches, that may be expected.
# More generally, the routines below can compile _all_ result cases
# found in the specified model subversion (VNN) directory and runset (VNNr0M) subdirectory

AEOyear <- "A2023"
run_years <- paste0("YR", c(2023, 2024, 2025, 2026, 2030, 2035, 2040, 2045, 2050))
# run_years <- paste0("YR", c(2023:2050))
default_switchstring <- "DE1CO1PC1"

runset_num <- 1 # subset of runs, expected in a subdirectory "r0$runset_num"

```

```{r defFn_generate_expected_case_names}
# add caseset case-number extension (N1..Nn) to filename
#  Unfortunately, this number is particular to the order in
#  which the cases are run, rather than their content
generate_expected_case_names <- function(curr_model_versionyear, 
                                         curr_model_subversion,
                                         curr_runsets,
                                         curr_runset_num,
                                         curr_AEOyear, 
                                         curr_run_years) {
  case_file_firstnames <- paste0(
    curr_model_versionyear, curr_model_subversion,
    curr_runsets[curr_runset_num], "_", curr_AEOyear, curr_run_years,
    default_switchstring
  )
  for (n in 1:length(case_file_firstnames)) {
    case_file_firstnames[n] <- paste0(case_file_firstnames[n], "SN", n)
  }
  return(case_file_firstnames)
}
```


```{r}
case_file_firstnames <- generate_expected_case_names(model_versionyear,
                             model_subversion,
                             runsets, runset_num,
                             AEOyear,
                             run_years)

case_file_firstnames
```

```{r}
# # note that tradition is upper-case v here for version number in directory names
input_subdirs <- c(paste0(toupper(model_subversion), runsets))
input_subdirs
```


```{r}
input_path_components <- c(inputdata_subpath, toupper(model_subversion), input_subdirs[runset_num])
# input_path_components
```

```{r findingDirs}
list_of_dirs <- list.dirs(here(inputdata_subpath), recursive = FALSE) # list of subdirectories to "inputdata_subpath"
list_of_dirs
```


```{r findingFiles}
# list of files in one runset directory
list_of_available_files <- list.files(here(paste(input_path_components, collapse = "/")), recursive = FALSE)
list_of_available_files
```

### Function Definitions

```{r defFn_str_cmp_num}
# following is currently unused
str_cmp_num <- function(s1, s2) {
  # Given 2 input strings s1, s2 of EQUAL LENGTH
  # returns position of first character difference
  #
  # (mapply is faster than loop, but only works for equal-length objects)
  mapply(
    function(x, y) which(x != y)[1],
    strsplit(s1, ""), strsplit(s2, "")
  )
}

# Test this fn
# str_cmp_num("Hello", "Helmo") == 4
```

```{r defFn_construct_case_file_firstname}

construct_case_file_firstname <- function(curr_model_versionyear,
                                          curr_model_subversion,
                                          curr_runsets,
                                          curr_runset_num,
                                          ncase,
                                          curr_switchstring) {
  # Construct the name of an expected file, for a number of setting and numberings
  # curr_runset_num: number of set of cases (usually years for same param values)
  # ncase: number of case within set of cases"
  # Globals: that define cases:
  #   AEOyear, run_years
  # Uses passed var `curr_switchstring instead of global `default_switchstring`
  c_file_name <- paste0(
    curr_model_versionyear,
    curr_model_subversion,
    curr_runsets[curr_runset_num], "_",
    AEOyear,
    run_years[ncase],
    curr_switchstring,
    "SN", ncase
  )
  c_file_name
}
```

```{r defFn_parse_case_file_firstname}
# ***unused ***
parse_case_file_firstname <- function(c_file_name) {
  # De-construct the name of an standard results file into identifying substrings
  case_name_pieces <- list(
    model_versionyear = str_sub(c_file_name, 1, 7),
    model_subversion = str_sub(c_file_name, 8, 10),
    runset = str_sub(c_file_name, 11, 13),
    AEOyear = str_sub(c_file_name, 14, 19),
    run_year = str_sub(c_file_name, 20, 25),
    switchstring = str_sub(c_file_name, 26, 34),
    ncase = str_sub(c_file_name, 35,) # rest of the string (not sure hom many chars)
  )
  return(case_name_pieces)
}

# example_file_name <- "OIP2023v32r03_A2023YR2050DE1CO1PC1SN9.xlsx"
# parse_case_file_firstname(example_file_name)
```

```{r defFn_read_one_excel_case}
read_one_excel_case <- function(case_file_firstname, path_spec) {
  # given a case file firstname, and path,
  # If found, attach the xlsx filetype and read the full/raw file.
  # Extract the essential table of data on statistics and percentiles for premium components "pi_*".
  # Return as clean df.
  full_case_path_spec <- here(
    paste(path_spec, collapse = "/"),
    paste0(case_file_firstname, ".xlsx")
  )

  # check for file existence first
  if (file.exists(full_case_path_spec)) {
    rslt <- read_excel(full_case_path_spec, range = "B11:AA39") %>% # table of variables by stat and percentiles
      rename(Year = `Year/Index`) %>% # clean up
      select(Year, Variable, starts_with("pi_")) %>% # keep pi components (drops derivatives, elasticities, case selector indices)
      filter(!is.na(Year) & Year != 0) # drop some stats not needed or complete (drops Std. Dev, Variance)
  } else {
    rslt <- NULL
  }
  return(rslt)
}
```


```{r defFn_extract_basic_rslts}
# ***superceded and unused***
extract_basic_rslts <- function(raw_rstl) {
  # from `raw_rstl`, the raw data read from a result xlsx file,
  # extract the essential table of data on statistics and percentiles for premium components "pi_*"
  rsltx <- rslt[10:38, 2:27] # "B11:AA39
  names(rsltx) <- rsltx[1, ]
  rsltx <- rsltx[2:nrow(rsltx), ]
  rsltx <- rsltx %>%
    rename(Year = `Year/Index`) %>%
    select(Year, Variable, starts_with("pi_")) %>%
    filter(!is.na(Year) & Year != 0)
}
```


```{r defFn_process_one_expected_set_of_cases}
process_one_expected_set_of_cases <- function(curr_model_versionyear, 
                                              curr_model_subversion, 
                                              curr_runsets,
                                              curr_runset_num,
                                              curr_switchstring) {
  # processes a set of cases based on _expected case names and switch settings
  #  The expected file names are constructed from anticipated year, numberings, and switch
  #  (A complication is that the run_num is fairly arbitrary, assigned based on order of execution.)
  print(paste0(
    "Processing `model_versionyear` ", curr_model_versionyear,
    ", `sub_version` ", curr_model_subversion
  ))
  rslts <- tibble() # initialize results df

  for (run_num in 1:9) { # assume 9 cases (years) per set.

    case_ffname <- construct_case_file_firstname(curr_model_versionyear,
                                          curr_model_subversion,
                                          curr_runsets,
                                          curr_runset_num,
                                          run_num,
                                          curr_switchstring)

    print(paste0("  reading case_set ", curr_runset_num, " case ", run_num, ", ", case_ffname))
    rslt <- read_one_excel_case(case_ffname, input_path_components)

    if (is.null(rslt)) {
      print(paste0("    Error or file not found for case ", case_ffname))
    } else { # non-NULL return from file read
      #  extract_basic_rslts() -> rslt
      if (nrow(rslts) == 0) { # first non-NULL result, initialize df format
        rslts <- rslt
      } else { # concatenate rows (expecting consistent columns)
        rslts <- bind_rows(rslts, rslt)
      }
    }
  }

  rslts <- rslts %>%
    mutate( # save run ID information with data
      model_versionyear = curr_model_versionyear,
      model_subversion = curr_model_subversion,
      runset_num = curr_runset_num,
      run_num = run_num,
      case_file_firstname = case_ffname
    )
  rslts
}
```

```{r defFn_process_one_found_set_of_cases}
process_one_found_set_of_cases <- function(mod_versionyear, mod_subversion, runset_n) {
  # processes a set of OIP case results found in a directory determined by mod_subversion and runset_n
  #  (no restriction on set of cases other than the follow conventions
  #    for start of the name, and directory/subdirectory)
  #
  #  `mod_versionyear`: typically  OIPYYYY, first part of model, base on year of analysis
  #  `mod_subversion`: is typically "Vnn", for some digits nn, a version of the OIP model
  #  `runset_n`: a single digit (number) indicating a set of cases completed (usually encoded in the OIP model name)
  #  # or? runset: a numerical vector of 1 or more numbers sets of runs (number typically are single digit)
  #
  #  The directory with the runs is expected to be:
  #     `paste0(toupper(mod_subversion),"/", toupper(mod_subversion), "r0", runset_num)`
  #  The expected file names are of xlsx type, and
  #  begin with OIPYYYYvNNrMM_" (the underscore distinguishing result cases from the model file itself
  # References Global: `inputdata_subpath`
  curr_runset <- paste0("r0", runset_n)
  input_dir <- toupper(mod_subversion)
  input_subdir <- paste0(toupper(mod_subversion), curr_runset)
  # these are the pieces of the path to the result file subdirectory, relative to the project directory
  input_path_compons <- c(inputdata_subpath, input_dir, input_subdir)

  # list of files in one runset directory
  list_of_rslt_files <- list.files(here(paste(input_path_compons, collapse = "/")), recursive = FALSE)
  expected_filename_start <- paste0(mod_versionyear, mod_subversion, curr_runset, "_")
  # filter out filenames that don't have expected initial characters for result files, or have std Excel extension
  list_of_rslt_files <- list_of_rslt_files[str_starts(list_of_rslt_files, expected_filename_start)]
  list_of_rslt_files <- str_sort(list_of_rslt_files[str_ends(list_of_rslt_files, ".xlsx")])

  print(paste0(
    "Processing `model_versionyear` ", mod_versionyear,
    ", `sub_version` ", mod_subversion, ", ", length(list_of_rslt_files), " cases:"
  ))
  rslts <- tibble() # initialize results df

  if (length(list_of_rslt_files) > 0) {
    for (case_file_name in list_of_rslt_files) { # assume 9 cases (years) per set.

      case_ffname <- str_remove(case_file_name, ".xlsx") # strip the file extension to just get file firstname
      print(paste0("  reading case_set ", runset_n, " case ", case_ffname))

      rslt <- read_one_excel_case(case_ffname, input_path_compons)

      if (is.null(rslt)) {
        print(paste0("    Error or file not found for case ", case_ffname))
      } else { # non-NULL return from file read
        #  extract_basic_rslts() -> rslt
        rslt <- rslt %>%
          mutate(
            run_num = NA,
            case_file_firstname = case_ffname
          )

        if (nrow(rslts) == 0) { # first non-NULL result, initialize df format
          rslts <- rslt
        } else { # concatenate rows (expecting consistent columns)
          rslts <- bind_rows(rslts, rslt)
        }
      }
    }

    rslts <- rslts %>%
      mutate( # these attribute apply to all result cases in this set
        model_versionyear = mod_versionyear,
        model_subversion = mod_subversion,
        runset_num = runset_n
      )
  } # if length(list_of_rslt_files) > 0
  rslts
}
```


### Execution Section

Test reading one result case

```{r}
testing <- F
if (testing) {
  runset_num <- 1
  run_num <- 1
  
  case_file_firstname <- construct_case_file_firstname(model_versionyear,
                                            model_subversion,
                                            runsets,
                                            runset_num,
                                            run_num,
                                            default_switchstring)
  rslt <- read_one_excel_case(case_file_firstname, input_path_components)
}
```

Test reading a set of cases based on expected names

```{r include=F, eval=F}
testing <- F
if (testing) {
  set_rslts <- process_one_expected_set_of_cases(model_versionyear,
                                                 model_subversion,
                                                 runsets,
                                                 runset_num,
                                                 default_switchstring)
    
  # saveExpectedRsltsToCSV
  if (save_results_to_CSV) {
    write_csv(set_rslts, here(output_subpath, paste0(model_versionyear, model_subversion, runsets[runset_num], ".csv")))
  }
  
  dim(set_rslts)
}
```

## Process Multiple OIP Result Cases

Read Multiple Sets of Cases, Taking any Result Files Found, and Store in CSV files

```{r defFn_process_all_result_cases}
process_all_result_cases <- function(curr_model_versionyears,
                                curr_model_subversions) {
  set_rslts_ls <- list()

  # grand Loop Over Models And Cases
  for (i in 1:2) { # could actually find the subdirectories, but whatever
    model_versionyear <- curr_model_versionyears[i]
    model_subversion <- curr_model_subversions[i]
    for (runset_num in 1:4) {
      listkey <- paste0(model_versionyear, model_subversion, "r0", runset_num)
      set_rslts_ls[[listkey]] <- process_one_found_set_of_cases(model_versionyear, model_subversion, runset_num)
  
      if (save_results_to_CSV) {
        write_csv(
          set_rslts_ls[[listkey]],
          here(output_subpath, paste0(model_versionyear, model_subversion, runsets[runset_num], ".csv"))
        )
      }
    } # for runset_num
  } # for i
  
  # collapse list of dfs to one df
  rslts <- bind_rows(set_rslts_ls) # shoule work even if some rslts df are empty
  
  return(rslts)
}
```

```{r}
all_rslts <- process_all_result_cases(model_versionyears,
                                      model_subversions)
```


```{r saveFounRsltsToCSV}
if (save_results_to_CSV) {
  write_csv(
    all_rslts,
    here(output_subpath, paste0("all_rslts", ".csv"))
  )
}
```

### Next Step: Run `oip_postprocess_pt2.Rmd`
